### 2018-02-09 Sections/cryptowars.tex
crypto

\comment{Add more advanced protocols? Decoy states, untrusted devices etc. etc.}

@@

Today there are countless freely available private-key cryptographic protocols available online, and some have been standardised by standards institutes. Currently, the Advanced Encryption Standard (AES) is a standard endorsed by the US government. AES is a block cipher, meaning that it divides data into small blocks of bits, each of which are encrypted independently, and operates with key lengths of up to 256 bits, making it very robust against brute-force attacks (Sec.~\ref{sec:brute_force_attack}).

@@

\subsection{Attacks on classical cryptography}

@@

The most obvious approach to cracking a crypto-system is to systematically try out all possible keys until we find one that correctly decodes the encrypted message. This is also the most na\"ive approach, and one which is computationally intractable for real-world key lengths. Specifically, for a key length of $k$ bits (\mbox{$k=256$} for AES), there are $2^k$ possible keys to try, and on average we will wait for $2^{k-1}$ trials before choosing the right one. Clearly an average waiting time of $2^{128}$ is not plausible!

@@

@@

@@

\comment{Note that when performing a brute-force attack against a private encryption key\index{Private-key encryption}, a quadratic speedup effectively halves the key length in terms of algorithmic runtime. Thus, in the quantum era private key lengths will need to be doubled.}

>>>>

\comment{Add more advanced protocols? CV? Decoy states, untrusted devices etc. etc.}

@@

Today there are countless freely available private-key cryptographic protocols available online, and some have been standardised by standards institutes. Currently, the Advanced Encryption Standard (AES) is a standard endorsed by the US government. AES is a block cipher, meaning that it divides data into small blocks of bits, each of which are encrypted independently, and operates with key lengths of up to 256 bits, making it very robust against brute-force attacks (Sec.~\ref{sec:attacks_on_class}).

@@

\subsection{Attacks on classical cryptography}\index{Cryptographic attacks}\label{sec:attacks_on_class}

@@

The most obvious approach to cracking a crypto-system is to systematically try out all possible keys until we find one that correctly decodes the encrypted message. This is also the most na\"ive approach, and one which is computationally intractable for real-world key lengths. Specifically, for a key length of $k$ bits (\mbox{$k=256$} for AES), there are $2^k$ possible keys to try, and on average we will wait for $2^{k-1}$ trials before choosing the right one. Clearly an average waiting time of $2^{255}$ is not plausible!

@@

Far better than waiting the age of the universe for the right key to turn up, is \textit{cryptanalysis}. Here we study patterns in input and output strings from a cipher utilising a particular key. There are many variations on this, but include techniques such as:
\begin{itemize}
	\item Known plaintext attacks: \comment{To do}
	\item Linear cryptanalysis: \comment{To do}
	\item Differential cryptanalysis: \comment{To do}
\end{itemize}

@@

In the case of RSA encryption, whose security derives from the believed computational hardness of factorising large numbers, the most efficient known classical algorithm for integer factorisation is the general number field sieve (GNFS), which has time-complexity,
\begin{align}
	O(\mathrm{exp}(c (\log n)^{\frac{1}{3}}) (\log\log n)^{\frac{2}{3}}),
\end{align}
which scales poorly for large $n$, keeping in mind that present-day implementations of RSA accommodate key lengths of up to 4,096 bits. \comment{Check this number}

@@

Note that when performing a brute-force attack against a private encryption key\index{Private-key encryption}, a quadratic speedup effectively halves the key length in terms of algorithmic runtime. Thus, in the quantum era private key lengths will need to be doubled to maintain the same security against brute-force attacks.
